---
title: "CRZ_311 noise complaints- descriptives"
author: "Nicole Itzkowitz"
date: "2026-01-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## set up
libraries
```{r, results='hide', message = 'hide'}
library(readr)
library(dplyr)
library(lubridate)
library(janitor)
library(ggplot2)
library(data.table)
library(sf)
library(sandwich)
library(lmtest)
library(nycOpenData)
```

311 data: January 01 2024 12:00 - January 01 2026 12:00, noise complaints only 
```{r data noise, results='hide'}
noise <- fread("/Users/ni2267/Documents/BEH/CRZ/CRZ_311/data/311_Service_Requests_20240101_20260101_noise.csv") |> clean_names()
# n=1,586,276

# alternative- pull using api but it doesn't work very well for large files- will come back to this
# noise <- nyc_pull_dataset(
#   dataset = "erm2-nwe9",
#   from = "2024-01-01",
#   to = "2025-12-31",
#   date_field = "created_date",
#   where = "complaint_type LIKE '%Noise%'",
#   limit = 200000,
#   timeout_sec = 1000)
                          
# remove complaints with missing values for long/lat (n = 6,810)
noise <- dplyr::filter(noise, !is.na(noise$longitude))

# n = 1,579,466 complaints (across all boroughs, all types of noise complaints)
```

geofence data
```{r data geo}
# CRZ geofence from:  https://data.ny.gov/Transportation/MTA-Central-Business-District-Geofence-Beginning-J/srxy-5nxn/about_data
crz <- read_sf("/Users/ni2267/Documents/BEH/CRZ/CRZ_311/data/MTA Central Business District Geofence_ Beginning June 2024_20250619/geo_export_d6c2fec2-3927-49c2-98c0-6c2b3edffa39.shp")
crz <- st_transform(crz, crs = "WGS84")

# ctrl area created in QGIS
ctrl <- read_sf("/Users/ni2267/Documents/BEH/CRZ/CRZ_311/data/crz_control/crz_control.shp")
ctrl <- st_transform(ctrl, crs = "WGS84")
```
note: eventually we need to update these to exclude some defined area along the border of the zones

dates
```{r data management clean dates}
noise$created_date <- lubridate::mdy_hms(noise$created_date)
noise$month_year <- format(noise$created_date, "%Y-%m")
```

classify traffic noise 
```{r data management classify traffic noise}
# check types of noise complaints
noise |> with(table(problem_detail_formerly_descriptor))

noise$traffic_noise <- case_when(noise$problem_detail_formerly_descriptor == "Car/Truck Horn" ~ 1,
                              noise$problem_detail_formerly_descriptor == "Car/Truck Music" ~ 1,
                              noise$problem_detail_formerly_descriptor == "Engine Idling" ~ 1,
                              noise$problem_detail_formerly_descriptor == "Noise: Vehicle (NR2)" ~ 1,
                              .default = 0)

noise |> with(table(problem_detail_formerly_descriptor, traffic_noise))

# n = 93,627 traffic related  noise complaints 

trafficnoise <- filter(noise, traffic_noise == 1)
```

geocode noise complaints
```{r data management zones}
# convert crashes into a shapefile
trafficnoise_sf <- st_as_sf(trafficnoise,
                       coords = c("longitude", "latitude"),
                       crs = "WGS84")

# attribute crash points to within CRZ
within_crz <- apply(st_within(trafficnoise_sf, crz, sparse = FALSE), 1, any)
trafficnoise_sf$crz <- as.integer(within_crz)

# attribute crash points to control zone
within_ctrl <- apply(st_within(trafficnoise_sf, ctrl, sparse = FALSE), 1, any)
trafficnoise_sf$ctrl <- as.integer(within_ctrl)

table(trafficnoise_sf$crz)
table(trafficnoise_sf$ctrl)

# crz_ctrl variable
trafficnoise_sf$crz_ctrl <- ifelse(trafficnoise_sf$crz == 1, 1,
                              ifelse(trafficnoise_sf$ctrl == 1, 0, NA))

trafficnoise_sf <- select(trafficnoise_sf, -c("crz", "ctrl"))

## subset data to only crz and ctrl noise complaints
df <- dplyr::filter(trafficnoise_sf, trafficnoise_sf$crz_ctrl == 0 | trafficnoise_sf$crz_ctrl == 1)
# n = 20,920 ALL traffic-related noise complaints within crz and ctrl zones

# df: each row represents a noise complaint , and crz_ctrl represents whether it was in  the crz or not
```

```{r data management zone month}
# create month variable
df$month_year <- paste0(year(df$created_date), "-", sprintf("%02d",month(df$created_date)))
df$pre_post <- ifelse(df$created_date > as.Date("2025-01-01"), 1, 0)
```

```{r}
# save shapefile, if needed for viz later on
df_subset <- select(df, c(unique_key, problem_formerly_complaint_type, crz_ctrl, traffic_noise, pre_post, month_year, geometry))
st_write(df_subset, "/Users/ni2267/Documents/BEH/CRZ/CRZ_311/data/crz_ctrl_noise_complaints.gpkg", append = F)
```

```{r data management aggregate by month year}
# aggregate by month year
zone_month_counts <- df %>%
  group_by(month_year, crz_ctrl, pre_post) %>%
  summarise(
    count_total   = n(),
    count_horn = sum(problem_detail_formerly_descriptor == "Car/Truck Horn", na.rm = TRUE),
    count_music = sum(problem_detail_formerly_descriptor == "Car/Truck Music", na.rm = TRUE),
    count_idle = sum(problem_detail_formerly_descriptor == "Engine Idling", na.rm = TRUE),
    count_horn_idle = count_horn + count_idle,
    .groups = "drop"
  )

zone_month_counts <- as.data.frame(zone_month_counts)
zone_month_counts <- select(zone_month_counts, -"geometry")
zone_month_counts$zone <- as.factor(ifelse(zone_month_counts$crz_ctrl == 1, "crz", "ctrl"))
```

## did: all road traffic-related noise complaints
### did plot
```{r did plot zone months}
summary(zone_month_counts$count_total)

zone_month_counts %>% 
  group_by(month_year, zone) %>%

  ggplot(aes(x = month_year, y = count_total, color = zone, group = zone)) +
  geom_line() +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 


# save image as pdf
# ggsave("did_month_plot.pdf", plot = fig1, device = "pdf", height = 6, width = 8, units = "in")

```

### did analysis
```{r did zone months}
did_zone_month <- glm(count_total ~ pre_post + crz_ctrl + pre_post*crz_ctrl, data = zone_month_counts, family = "poisson")

summary(did_zone_month)
exp(coef(did_zone_month))
exp(confint(did_zone_month))

#cluster-robust standard errors
clustered_vcov <- vcovCL(did_zone_month, cluster = ~month_year)
exp(coef(coeftest(did_zone_month, vcov = clustered_vcov)))
exp(confint(coeftest(did_zone_month, vcov = clustered_vcov)))
```

## did: congestion related noise complaints only
restricting to ONLY complaints about car/truck horn and engine idling

### did2 plot
```{r did2 plot zone months}
summary(zone_month_counts$count_horn_idle)

zone_month_counts %>% 
  group_by(month_year, zone) %>%

  ggplot(aes(x = month_year, y = count_horn_idle, color = zone, group = zone)) +
  geom_line(linewidth = 1.2) +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "dashed", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) -> did2

print(did2)

ggsave(
  "/Users/ni2267/Documents/BEH/CRZ/CRZ_311/output/crz311.png",
  did2,
  width = 7,
  height = 5,
  dpi = 300
)

# also save to icben presentation- can delete this later
ggsave(
  "/Users/ni2267/Documents/Columbia/Conferences/ICBEN_2026/presentation/crz311.png",
  did2,
  width = 7,
  height = 5,
  dpi = 300
)
```
### descriptives
```{r}
# pre-intervention mean
zone_month_counts |> 
  filter(pre_post == 0) |> 
  group_by(crz_ctrl) %>%
    summarise(
      mean_val = mean(count_horn_idle, na.rm = TRUE)
  )

# post-intervention mean
zone_month_counts |> 
  filter(pre_post == 1) |> 
  group_by(crz_ctrl) %>%
    summarise(
      mean_val = mean(count_horn_idle, na.rm = TRUE)
  )
```

### did2 analysis
```{r did2 zone months}
did_zone_month2 <- glm(count_horn_idle ~ pre_post + crz_ctrl + pre_post*crz_ctrl, data = zone_month_counts, family = "poisson")

summary(did_zone_month2)
exp(coef(did_zone_month2))
exp(confint(did_zone_month2))

#cluster-robust standard errors
clustered_vcov <- vcovCL(did_zone_month2, cluster = ~month_year)
exp(coef(coeftest(did_zone_month2, vcov = clustered_vcov)))
exp(confint(coeftest(did_zone_month2, vcov = clustered_vcov)))
```


## did: congestion related noise complaints only v3
restricting to ONLY complaints about car/truck horn 

### did3 plot
```{r did3 plot zone months}
summary(zone_month_counts$count_horn)

zone_month_counts %>% 
  group_by(month_year, zone) %>%

  ggplot(aes(x = month_year, y = count_horn, color = zone, group = zone)) +
  geom_line() +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

### did3 analysis
```{r did3 zone months}
did_zone_month3 <- glm(count_horn ~ pre_post + crz_ctrl + pre_post*crz_ctrl, data = zone_month_counts, family = "poisson")

summary(did_zone_month3)
exp(coef(did_zone_month3))
exp(confint(did_zone_month3))

#cluster-robust standard errors
clustered_vcov <- vcovCL(did_zone_month3, cluster = ~month_year)
exp(coef(coeftest(did_zone_month3, vcov = clustered_vcov)))
exp(confint(coeftest(did_zone_month3, vcov = clustered_vcov)))
```


## did: congestion related noise complaints only v4
restricting to ONLY complaints about engine idling 

### did4 plot
```{r did4 plot zone months}
summary(zone_month_counts$count_idle)

zone_month_counts %>% 
  group_by(month_year, zone) %>%

  ggplot(aes(x = month_year, y = count_idle, color = zone, group = zone)) +
  geom_line() +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

### did4 analysis
```{r did4 zone months}
did_zone_month4 <- glm(count_idle ~ pre_post + crz_ctrl + pre_post*crz_ctrl, data = zone_month_counts, family = "poisson")

summary(did_zone_month4)
exp(coef(did_zone_month4))
exp(confint(did_zone_month4))

#cluster-robust standard errors
clustered_vcov <- vcovCL(did_zone_month4, cluster = ~month_year)
exp(coef(coeftest(did_zone_month4, vcov = clustered_vcov)))
exp(confint(coeftest(did_zone_month4, vcov = clustered_vcov)))
```

## separate ITS for each zone- traffic noise 

```{r its set up}
# create season variable, by month
# winter: dec, jan, feb
# spring: mar, apr, may
# summer: jun, jul, aug
# autumn: sept, oct, nov
zone_month_counts$season <- ifelse(grepl("-01|-02|-12", zone_month_counts$month_year), 1, 
                                  ifelse(grepl("-03|-04|-05", zone_month_counts$month_year), 2, 
                                         ifelse(grepl("-06|-07|-08", zone_month_counts$month_year), 3, 4)))

```

### its: plots, all complaints
#### crz all
```{r its crz all plot}
zone_month_counts %>% 
  filter(crz_ctrl == 1) |> 
   ggplot(aes(x = month_year, y = count_total, color = zone, group = zone)) +
  geom_line( color = "#F8766D") +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

#### ctrl all
```{r its ctrl all plot}
zone_month_counts %>% 
  filter(crz_ctrl == 0) |> 
   ggplot(aes(x = month_year, y = count_total, color = zone, group = zone)) +
  geom_line(color = "#00BFC4") +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

###: its: analysis, all complaints
#### crz: all
```{r its: crz}
zone_month_counts %>%
  filter(crz_ctrl == 1) %>%
  mutate(time = row_number()) %>%
glm(count_total ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> crz_month_season

summary(crz_month_season)
exp(coef(crz_month_season))
exp(confint(crz_month_season))
```

#### ctrl: all
```{r its: ctrl}
zone_month_counts %>%
  filter(crz_ctrl == 0) %>%
  mutate(time = row_number()) %>%
glm(count_total ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> ctrl_month_season

summary(ctrl_month_season)
exp(coef(ctrl_month_season))
exp(confint(ctrl_month_season))
```

### its: horn and idling only
#### crz horn and idling
```{r its crz horn and idling plot}
zone_month_counts %>% 
  filter(crz_ctrl == 1) |> 
   ggplot(aes(x = month_year, y = count_horn_idle, color = zone, group = zone)) +
  geom_line( color = "#F8766D") +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 
```

#### ctrl horn and idling
```{r its ctrl horn and idling plot}
zone_month_counts %>% 
  filter(crz_ctrl == 0) |> 
   ggplot(aes(x = month_year, y = count_horn_idle, color = zone, group = zone)) +
  geom_line(color = "#00BFC4") +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1))
```
#### crz: horn and idling
```{r its: crz horn and idling}
zone_month_counts %>%
  filter(crz_ctrl == 1) %>%
  mutate(time = row_number()) %>%
glm(count_horn_idle ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> crz_month_season

summary(crz_month_season)
exp(coef(crz_month_season))
exp(confint(crz_month_season))
```

#### ctrl: horn and idling
```{r its: ctrl horn and idling}
zone_month_counts %>%
  filter(crz_ctrl == 0) %>%
  mutate(time = row_number()) %>%
glm(count_horn_idle ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> ctrl_month_season

summary(ctrl_month_season)
exp(coef(ctrl_month_season))
exp(confint(ctrl_month_season))
```


## negative controls: residential noise

### data set up 
311 data: January 01 2024 12:00 - January 01 2026 12:00, noise complaints only 
```{r data residential noise, results='hide'}
res_noise <- filter(noise, problem_formerly_complaint_type == "Noise - Residential")
# n=841,185
```

```{r data geocode residential noise}
# convert crashes into a shapefile
resnoise_sf <- st_as_sf(res_noise,
                       coords = c("longitude", "latitude"),
                       crs = "WGS84")

# attribute crash points to within CRZ
within_crz <- apply(st_within(resnoise_sf, crz, sparse = FALSE), 1, any)
resnoise_sf$crz <- as.integer(within_crz)

# attribute crash points to control zone
within_ctrl <- apply(st_within(resnoise_sf, ctrl, sparse = FALSE), 1, any)
resnoise_sf$ctrl <- as.integer(within_ctrl)

# crz_ctrl variable
resnoise_sf$crz_ctrl <- ifelse(resnoise_sf$crz == 1, 1,
                              ifelse(resnoise_sf$ctrl == 1, 0, NA))

resnoise_sf <- select(resnoise_sf, -c("crz", "ctrl"))

## subset data to only crz and ctrl noise complaints
res_df <- dplyr::filter(resnoise_sf, resnoise_sf$crz_ctrl == 0 | resnoise_sf$crz_ctrl == 1)
# n = 126,330 ALL residential noise complaints within crz and ctrl zones
```

```{r data check residential complaint type}
table(res_df$problem_detail_formerly_descriptor)
```

```{r data management residential zone month}
# create month variable
res_df$month_year <- paste0(year(res_df$created_date), "-", sprintf("%02d",month(res_df$created_date)))
res_df$pre_post <- ifelse(res_df$created_date > as.Date("2025-01-01"), 1, 0)
```

```{r data management residential aggregate by month year}
# aggregate by month year
res_zone_month_counts <- res_df %>%
  group_by(month_year, crz_ctrl, pre_post) %>%
  summarise(
    count_total   = n(),
    count_bang = sum(problem_detail_formerly_descriptor == "Banging/Pounding", na.rm = TRUE),
    count_music = sum(problem_detail_formerly_descriptor == "Loud Music/Party", na.rm = TRUE),
    count_talk = sum(problem_detail_formerly_descriptor == "Loud Talking", na.rm = TRUE),
    count_tv = sum(problem_detail_formerly_descriptor == "Loud Television", na.rm = TRUE),
    .groups = "drop"
  )

res_zone_month_counts <- as.data.frame(res_zone_month_counts)
res_zone_month_counts <- select(res_zone_month_counts, -"geometry")
res_zone_month_counts$zone <- as.factor(ifelse(res_zone_month_counts$crz_ctrl == 1, "crz", "ctrl"))
```

### negative control did
```{r did plot residential zone months}
#summary(res_zone_month_counts$count_total)

res_zone_month_counts %>% 
  group_by(month_year, zone) %>%

  ggplot(aes(x = month_year, y = count_tv, color = zone, group = zone)) +
  geom_line() +
  labs(x = "month", y = "zone count", color = "CRZ") +
  scale_x_discrete(breaks = function(x) x[seq(1, length(x), by = 2)]) +
  geom_vline(xintercept = "2025-01", linetype = "solid", color = "black") +
  theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, hjust = 1)) 

```

```{r did residential zone months}
did_res_zone_month <- glm(count_bang ~ pre_post + crz_ctrl + pre_post*crz_ctrl, data = res_zone_month_counts, family = "poisson")

summary(did_res_zone_month)
exp(coef(did_res_zone_month))
exp(confint(did_res_zone_month))

#cluster-robust standard errors
clustered_vcov <- vcovCL(did_zone_month2, cluster = ~month_year)
exp(coef(coeftest(did_zone_month2, vcov = clustered_vcov)))
exp(confint(coeftest(did_zone_month2, vcov = clustered_vcov)))
```

### separate its: negative control residential noise 
```{r its residential set up}
# create season variable, by month
# winter: dec, jan, feb
# spring: mar, apr, may
# summer: jun, jul, aug
# autumn: sept, oct, nov
res_zone_month_counts$season <- ifelse(grepl("-01|-02|-12", zone_month_counts$month_year), 1, 
                                  ifelse(grepl("-03|-04|-05", zone_month_counts$month_year), 2, 
                                         ifelse(grepl("-06|-07|-08", zone_month_counts$month_year), 3, 4)))

```
#### crz: its 
```{r its: crz residential noise}
res_zone_month_counts %>%
  filter(crz_ctrl == 1) %>%
  mutate(time = row_number()) %>%
glm(count_total ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> crz_res

summary(crz_res)
exp(coef(crz_res))
exp(confint(crz_res))
```

### ctrl: its 
```{r its: crz residential noise}
res_zone_month_counts %>%
  filter(crz_ctrl == 0) %>%
  mutate(time = row_number()) %>%
glm(count_total ~ time + pre_post + time*pre_post + season, data = ., family = "poisson") -> ctrl_res

summary(ctrl_res)
exp(coef(ctrl_res))
exp(confint(ctrl_res))
```